### 2020-08-27 writing/ch1.tex
starting revisions

\hypertarget{the-construction-of-space-in-ecological-models}{%
\section{The Construction of Space in Ecological
Models}\label{the-construction-of-space-in-ecological-models}}

\begin{quote}
The electron is a theory we use; it is so useful in understanding the
way nature works that we can almost call it real.

\begin{flushright}Richard P. Feynman\end{flushright}
\end{quote}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Life on Earth takes on an endless diversity of forms. Yet, the
characteristics that define `life' impose parameters on the properties
that biological entities must exhibit, and constraints on how these
entities can change over time. These are the fundamental principles of
evolution and ecology---entities that reproduce themselves more than
average become more frequent, there is a limit on resources, and so on.
Over the billions of years life has been on this planet, these forces
have produced an astonishing diversity of forms and functions. The
forces that drive evolutionary and ecological processes do not occur on
a single scale---they emerge out of the interactions that occur across
all levels of spatial and temporal biological organization
(@levin\_problem\_1992).

Understanding how ecological processes change across space, within and
across scales, is central to the questions of ecology and evolution. The
spatial distribution of individuals is central to behavioral ecology,
the spatial distribution of species to biogeography, the spatial
distribution of genes to speciation, and so on. Further, the influence
of space on ecological processes has significant applied importance as
human activity has drastically altered the face of our planet in an
extremely short period of time (@bigelow\_major\_2017,
@vitousek\_human\_1997). Land-use change has shifted the amount and
configuration of Earth's habitats, and climate change is shifting the
spatial distribution of environmental factors and species' distributions
(@pereira\_scenarios\_2010, @thomas\_climate\_2010,
@loarie\_velocity\_2009). This drastic change in the structure of our
planet's terrain has had overwhelmingly negative effects on the
biodiversity to which it is home (@haddad\_habitat\_2015,
@rands\_biodiversity\_2010). This has, in turn, degraded ecosystem
functioning, and the services provided by ecosystem functioning, around
the globe (@tilman\_biodiversity\_2014, @cardinale\_biodiversity\_2012).

Historically, studies of the effects of landscape structure on
biodiversity have fallen under the broad umbrella of ``habitat
fragmentation'' (@collinge\_ecology\_2009). However, there has been
significant debate regarding what is meant by this term. The primary
distinction here is the difference between the word \emph{fragmentation}
as it is commonly used as a catchall to describe a combination of
habitat loss and ensuing subdivision, and \emph{fragmentation per se},
which is purely the spatial separation or reconfiguration of the same
amount of habitat (@fahrig\_effects\_2003). The effect of spatial
subdivision independent of habitat loss (fragmentation \emph{per se}) on
biodiversity is still hotly contested (@fahrig\_is\_2019,
@fletcher\_is\_2018, @fahrig\_ecological\_2017,
@ewers\_confounding\_2006). However, it is abundantly clear that the
combined effects of habitat loss and subdivision have had negative
effects on the planet's biodiversity (@haddad\_habitat\_2015). As a
result, a major focus on conservation efforts has become understanding
and predicting the simultaneous effects of habitat loss and subdivision
on ecosystem structure and stability (@fletcher\_spatial\_2019).

In order to understand and predict the consequences of landscape change,
we first need models to bridge the gap between the data we can collect
and the inferences and predictions we wish to make about the world.
Before we can construct quantitative models to make predictions with
real data, we first need to conceptualize the process we are modeling in
space. Over the years, ecologists have considered many different ways of
representing the structure of space and its influence on ecological
processes, using a variety of frameworks and tools to represent
different processes across different scales. Here, we examine the role
that model construction plays in scientific epistemology to better
understand how models can be used and compared, and review the various
constructions of space ecological models have used in the past.

\hypertarget{what-is-a-model}{%
\subsection{What is a Model?}\label{what-is-a-model}}

Science is fundamentally a theory of epistemology---a way of knowing.
Within the constraints of scientific epistemology, for a scientific
theory to be considered valid, it must be falsifiable through
observation, and for a theory to be disproven through observation, it
must make predictions that may or may not agree with observed reality
(@popper\_logic\_1934). In this view, a scientific model can be seen as
a function---some ontological object, \(f\), that maps measurable input
conditions, \(\vec{x}\), to a predicted outcome, \(y=f(\vec{x})\).

Constructing models is inherently a creative act. Before we can
introduce any quantitative representation of a process, we need to
conceptualize what it is we should even be quantifying. In \emph{What is
Philosophy?}, Deleuze \& Guattari (1991) answer the eponymous question
by suggesting philosophy is the act of concept creation. They write:

\begin{quote}
In fact, the sciences, arts, and philosophies are all equally
creators\ldots{} Concepts do not wait for us ready-made, like celestial
bodies. There is no heaven for concepts. They must be invented,
fabricated, or rather created, and would be nothing without the
signature of those who create them. (Deleuze \& Guattari 1991)
\end{quote}

In this view, scienctific modeling, too, is an act of creation, not
discovery. The concepts we use to frame natural systems are not inherent
properties of nature. Especially in ecology, it often seems the atomic
unit is the individual, yet this ignores differences in that
individual's cells, driven by biochemical processes that are the product
of interactions molecules within those cells, and so on---there is not
natural or correct scale on which to model natural processes
(@levin\_problem\_1992), and the conceptual frameworks we use to model
ecosystems are tools created by the modeler. Just as Newtonian Gravity,
General Relativity, and the Principle of Least-Action all describe and
predict the same phenomena using entirely different conceptual
frameworks (@arkani-hamed\_future\_2012, @feynman\_feynman\_1965), valid
ecological models can be framed in a variety of different ways.

A model's ``correctness'' can be measured as the difference between its
prediction, \(y\), and observed reality, \(\hat{y}\)
(@jaynes\_probability\_2003). However, all models have information baked
into them based on how they define \(f\), and how much information is
already contained in \(\vec{x}\) . Further, as we add more parameters to
a model it becomes harder to disprove that our predictions \(y\) don't
agree with \(\hat{y}\). To deal with this general problem of
overfitting, we turn to the theory of model selection. Model selection
is rooted in information theory---many common methods for model
selection---Minimum Description Length, AIC, BIC---revolve around the
huerstic that better models maximize that ratio of how much information
a model provides in predictive power to how much is built information
into the model structure (@jaynes\_probability\_2003,
@stine\_model\_2004). Other methods, like crossvalidation, aim to test
the predictive accuracy of a model by testing it on data it hasn't
``seen'' yet. (@konishi\_information\_2008). Either way, this provides
criteria by which to assess how models compare to one another in their
capacity to explain natural phenomena. However, what is actually meant
by `model' in practice varies widely. We use models for a variety of
things---inference, hypothesis-testing, prediction---and for these
different purposes we use a different methods to construct \(f\).

Breiman (2001) partitions models into two types: data models and
algorithmic models. Both map input conditions to output conditions, but
the difference lies is in the approach we take to create \(f\).
Algorithmic models aim to construct \(f\) as an algorithm which
accurately predicts output conditions based on input conditions. The
algorithm tends to have little relation to the mechanisms producing the
data being modeled, but rather is designed with the aim of accurately
finding patterns in data---methods of this type tend to receive the
somewhat ambiguous label ``machine learning''. These models can be
exceptionally useful for prediction, but this comes at the cost that
they do not reliably provide information about the mechanisms producing
the data.

In contrast, data models attempts to model the way in which the data is
generated. We can further partition data models into statistical models
and process models. The vast majority of scientific studies use
statistical models (@breiman\_statistical\_2001), which conceptualize
\(f\) by assuming the output of \(f\) is some combination of its input
conditions, \(x_i\), each of which is drawn from an assumed distribution
for each observation. Models of this form have seen extensive use
because they are easily generalizable and can be applied to a variety of
problems. Further, by assuming that measured quantities follow
distributions that we can describe analytically, we are able to compose
likelihood functions which can then be used to fit models to data using
a variety of methods, both frequentist and Bayesian. In contrast,
process models aim to construct \(f\) to represent the hypothesized
mechanism that produces the data. This can be done in a variety of
methods---for example, \(f\) could be written as a stochastic
differential equation, as we will do in the next chapter, or \(f\) could
be described as a set of properties between agents in an
individual-based model. Process models can vary in the degree to which
they represent the fidelity of the actual process based on the scale and
detail of representation considered. However, this comes with the cost
that, in general, process models are much more difficult to fit data to
than statistical models. However, recent advances in Approximate
Bayesian Computation (@beaumont\_approximate\_2010) make it feasible to
fit data to complex simulation models which cannot be described by an
analytic likelihood function. This can give us a way to test between
mechanistic process models, given that we have a reasonably small number
of potential mechanisms we want to test.

Now, we turn to the history of spatial models in ecology to survey the
the ways in which ecological process models have conceptualized space.

\hypertarget{a-history-of-the-construction-of-space-in-ecology}{%
\subsection{A History of the Construction of Space in
Ecology}\label{a-history-of-the-construction-of-space-in-ecology}}

\hypertarget{the-theory-of-island-biogeography}{%
\subsubsection{The Theory of Island
Biogeography}\label{the-theory-of-island-biogeography}}

Nearly all of landscape ecology has its roots in the Theory of Island
Biogeography (TIBG, @macarthur\_theory\_1967). The TIBG modeled a system
of oceanic islands, each with a different size, corresponding to
resource levels. The primary motivation of the theory was to provide a
mechanistic explanation of the species area relationship, and so the
core of the TIBG relates the spatial structure of the islandscape---the
size of each island and the distances between them---to observed
patterns of species richness.

Although the TIBG was conceptualized for terrestrial communities on
oceanic islands, it quickly was applied elsewhere under the assumption
that many human-altered landscapes are well-approximated by an island
structure---isolated regions of homogeneous landscape separated by
inhabitable matrix (@haila\_conceptual\_2002, @macarthur\_theory\_1967,
@preston\_canonical\_1962). The analogy between islands and subdivided,
patchy landscapes was first made by Preston (1962), and is briefly
invoked in the TIBG. However, this analogy went on to strongly influence
how landscape ecologists modeled space in terrestrial landscapes in the
future. Traces of the central ideas of the TIBG---area as a proxy for
resource availability, richness as a product of an equilibrium between
colonization and extinction, that patches are internally
homogenous---can be seen all throughout the models of landscape ecology
that followed.

\hypertarget{metapopulation-theory}{%
\subsubsection{Metapopulation Theory}\label{metapopulation-theory}}

\hypertarget{classical-metapopulation-theory}{%
\paragraph{Classical Metapopulation
Theory}\label{classical-metapopulation-theory}}

The metapopulation framework was introduced by Levins (1969)---who
originally defined a metapopulation as `a population of populations'.
Metapopulation theory, like the TIBG, considers the dynamics of
occupancy due to colonization or extinction. Levins' model considers a
system of infinite populations, each with a fixed probability of
colonization, \(c\), or local extinction, \(e\) , each generation. In
this formulation, the proportion of populations occupied at a given
time, \(p\), is modeled as as

\[\frac{dp}{dt} = cp(1-p)-ep\]

From this, we can derive that the system will persist, i.e.~\(p \to 1\),
if and only if \(\frac{c}{e} >1\), and \(p\to0\) otherwise
(@levins\_demographic\_1969). The aim of this model was not to be used
to predict the consequences of subdivision in an empirical
system---rather, this is what Okubo (1978) calls a ``toy model'': a way
of examining the consequences imposed by an oversimplified version of
the dynamics in question. In reality, stochasticity in
colonization/extinction events cause real systems to diverge from the
deterministic predictions made by Levins' model, and the spatially
implicit structure of this model made it difficult to assess the
influence of landscape structure or use for prediction in a real
landscape.

\hypertarget{modern-metapopulation-theory}{%
\paragraph{Modern Metapopulation
Theory}\label{modern-metapopulation-theory}}

Metapopulation theory was more formally applied to fragmented landscapes
by Hanski and Ovaskainen (@hanski\_practical\_1994,
@hanski\_metapopulation\_2000), who adapted Levins' framework of a
network of occupied/unoccupied patches, and restricted it to a finite
number of populations with spatially-explicit locations, each with a
unique probability of colonization arising from the metapopulation's
spatial structure. This introduces more parameters---each population
\(i\) has a spatial coordinate in \(\mathbb{R}^2\), \((x_i, y_i)\), and
an area associated with it, \(A_i\). Here, like in the TIBG, area is
still a proxy for the resource availability in that patch, as the
probability of a patch going extinct in a generation is is inversely
proportional to its area, and directly proportional to its ability to
serve as a source for colonizers. Similarly, in a departure from Levins'
model, the probability of an unoccupied patch getting colonized during
any timestep is inversely proportional to how far away it is from
occupied patches. This relationship of isolation-by-distance (IBD),
called the dispersal kernel, was initially modeled as an exponential
(@hanski\_practical\_1994), \(f(d_{ij})= \ e^{-\alpha d_{ij}}\), which
has seen extensive use since.

What made Hanski and Ovaskainen's theory capable of interfacing with
real data is their use of incidence-function models (IFMs,
@hanski\_practical\_1994). The value of the incidence function \(J_i\),
for any population \(i\), is the long run proportion of time that
population \(i\) is occupied. From this, one can compute the probability
of changing from any state of occupancy to any other state, which
provides us with a likelihood function to estimate parameters from
occupancy data in real systems. Much like the TIBG, this is based on the
assumption that when the data we collect is indicative of a system that
has approached its long-term equilibrium state. This, however, has
enabled the extensive use of IFMs in conservation
(@risk\_robust-design\_2011, @macpherson\_metapopulation\_2011,
@vos\_incidence\_2000, @hanski\_quantitative\_1996). More recent
metapopulation studies use so-called Dynamical Models
(@fletcher\_spatial\_2019), which model occupancy in space as a
Bernoulli distribution, with weighted parameters to account for
isolation-by-distance/resistance, environmental variance, or other
covariates.

\hypertarget{sources-and-sinks}{%
\paragraph{Sources and Sinks}\label{sources-and-sinks}}

Another theoretical output of metapopulation theory is the conceptual
framework of sources and sinks. Source-sink metapopulations emerged from
the theory of birth-immigration-death-emigration (BIDE) demographic
models (@pulliam\_sources\_1988), the idea being that different
locations in space have differing levels of resource availability, and
that as a result, have different expected values of each of these
parameters every generation. Populations with more births than deaths
and more emigrants than immigrants are `sources', populations with
higher deaths than births are `sinks'. This conceptual framework has
since spread outside population ecology, and gone on to be influential
across metacommunity (@mouquet\_community\_2003), population genetics
(@gaggiotti\_population\_1996), and coevolutionary theory
(@thompson\_coevolution\_2002).

Within both the metapopulation framework and the TIBG, it is assumed
that there isn't heterogeneity within patches, and that the region
between habitats is entirely inhospitable. However, these assumptions
would later be challenged by the increasing availability of data that
allowed ecologists to think about space continuously.

\hypertarget{continuous-space-lattice-models-and-the-matrix}{%
\subsubsection{Continuous Space, Lattice Models, and the
Matrix}\label{continuous-space-lattice-models-and-the-matrix}}

One natural limit of the above models is that ecological processes do
not occur in discrete space. Although continuous-space models have long
been used in theoretical studies (@haldane\_theory\_1948,
@hastings\_global\_1978, Okubo 1978), the difficulty inherent to both
collecting `continuous' spatial data, plus the increased mathematical
complexity induced by considering processes in continuous space, long
made such models impractical to apply to an empirical system.

This changed due to two major developments: the increased availability
of remotely sensed land cover data and the onset of computational power
to feasibly run simulation models that well-approximate continuous
space. The advent of remote sensing imagery enabled construction of
global land-cover maps, which function as approximations of continuous
space on large scales. In response, landscape research focused on
studying various aspects of patch geometry in raster data
(@mcgarigal\_fragstats\_1995).

This, plus the continued growth of computational power, paved the way
for models using rasters to approximate continuous space, meaning we
were no longer limited in our ability to model continuous space with the
tools of analytical math. Lattice-based simulation models have been
applied to many questions, including the theory of percolation in
habitat loss (@bascompte\_habitat\_1996), evolutionary rescue models
(@boeye\_more\_2013), epidemiology models (@jeltsch\_pattern\_1997),
coevolutionary models (@sisterson\_coevolution\_2004), adaptive
radiation (@gavrilets\_dynamic\_2005), predator-prey models
(@matsuda\_statistical\_1992), and so on.

These developments led us to more complex conceptual models of landscape
structure. The relationship of different land cover types in relation to
one another marks a significant shift away from the habitat is either
suitable-or-unsuitable binary that has dominated landscape ecology for
many years, and to other models including the so-called mosaic model
which considers all elements of the matrix as important to understanding
landscape structure (Fortin \& Fletcher 2019, Wiens 1995).

\hypertarget{resistance-connectivity-and-spatial-graphs}{%
\subsubsection{Resistance, Connectivity, and Spatial
Graphs}\label{resistance-connectivity-and-spatial-graphs}}

The field of landscape genetics, introduced by Manel et al.~(2003), has
sought to synthesize an understanding of landscape structure with
population genetic models using the recent availability of
next-generation sequencing technology. This field has contributed many
conceptual frameworks to the study of space, including idea of
isolation-by-resistance (IBR, @mcrae\_isolation\_2006,
@mcrae\_using\_2008). IBR argues that different landscape features,
typically represented as different land cover types, impose different
strengths of ``resistance'' to dispersal or movement. Each cell in a
raster landscape, called a resistance surface, is modeled as a resistor,
and the minimum resistance path between any two points in the raster can
be estimated using circuit theory (@spear\_use\_2010). More recently
models of gene flow have been used to validate resistance surfaces
(@mateo-sanchez\_estimating\_2015), and construct them from from a
combination of genetic and landscape data
(@peterman\_resistancega\_2018).

The concept of landscape connectivity has seen much recent interest in
conservation ecology (@robertson\_isolating\_2018,
@mitchell\_linking\_2013, @carroll\_use\_2012,
@krosby\_ecological\_2010). The effect of connectivity on ecological
processes has been clear since the concept was first introduced
(@turner\_landscape\_1989, @taylor\_connectivity\_1993), however, the
theory of resistance surfaces has proven an invaluable tool in
quantifying connectivity (@kool\_population\_2013). The use of spatial
graphs in connectivity models has also seen recent interest, as they
provide a generalizable framework to study pattern and process in
landscapes across different scales (@chubaty\_r\_2020,
@dale\_graphs\_2010, @minor\_graph-theory\_2008,
@urban\_landscape\_2001).

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

Here, we have focused on the various ways that spatial models have been
constructed in ecology---however, as mentioned earlier, scientists use
models to serve a number of different functions, and different ways of
modeling space have different utility in different circumstances. In the
next chapter, we restrict our focus to simulation models. We suggest
that when planning efforts to restore landscape connectivity, there is
immense value in using simulation models to predict the effects of
potential corridors on functional connectivity
(@epperson\_utility\_2010).

\pagebreak